---
title: 'Marketing Analytics Project: Black Friday'
author: "Weicheng Zhang, Zilei Zhang, Maura Oray, Gloria Zheng"
output:
  word_document: default
  html_document:
    df_print: paged
---

**1. Introduction** 
**Objective:**
* Look at the distribution of customer demographics
* Determine which variables are highly correlated with purchase
* Identify which products are high-selling and the characteristics of their customers

The dataset comes from a competition hosted by Analytics Vidhya.

Group Github url: https://github.com/MauraO/Marketing-Analytics-Project
```{r results='hide',message=FALSE, warning=FALSE}
#Load libraries
library(VIM)
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(devtools)
devtools::install_github('cttobin/ggthemr')
library(ggthemr)

BF <- read_csv("BlackFriday.csv")
```
**2. Data Description**
**a. Describe the conceptual measure types of the different variables**

User_ID (discrete, nominal), Product_ID (discrete, nominal), Gender (discrete, nominal), Age (discrete, ordinal), Occupation (discrete, nominal), City_Category (discrete, nominal), Stay_in_Current_City_Years (continuous, ratio), Marital_Status (discrete, nominal), Product_Category_1 (discrete, nominal), Product_Category_2 (discrete, nominal), Product_Category_3 (discrete, nominal), and Purchase (continuous, ratio). 

**b. Mention all the steps you took to clean the data**
**Step 1. Check if the data contains missing values**
```{r results='hide',message=FALSE, warning=FALSE}
sum(complete.cases(BF))
aggr(BF,cex.axis = .4) #164,278 NA values
```
In the above visualization, we want to see which variables had the most NA values. We see that the NA values were only concentrated in Product Category 2 and Product Category 3 (red indicates the location of NA values). Because these two variables are sub category of Product Category 1 and our further analysis does not use them, we delete them and make a new data set later. 

**Step 2. Check the structure of the data**
```{r results='hide', message=FALSE, warning=FALSE}
summary(BF)
str(BF)
```
In this part, we want to see the type of variables for later use(group_by). We find all of them are not factor, therefore, we come to the next step.

**Step 3. Type coercion**
```{r results='hide',message=FALSE, warning=FALSE}
BF[1:11] = lapply(BF[1:11], factor)
```
The majority of the variables were either integers or characters. We decided to coerce the variables into factors (all except purchase) because the majority of them are categorical, and coercing to factors makes for easier and clearer analysis(group_by).

**Step 4. Tidy the data**
```{r results='hide',message=FALSE, warning=FALSE}
#Select demographic data
demo = BF %>% 
  select(-starts_with("Product")) %>%
  group_by(User_ID) %>%
  mutate(Purchase = sum(Purchase))
demo = demo[!duplicated(demo$User_ID), ]
```
Because we decide to focus more on customer analysis instead of the product and Product Category 2 and Product Category 3 have a lot of missing data, we delete all of the product category columns to make a clearer customer data set. Moreover, we find that the Customer IDs are repeated for many times which is redundant and we need a sum purchase number. Therefore, we use the customer all purchase to replace the individual per product purchase and delete the redundant customer IDs. 

**3. Summary statistics and Data Visualizations** 
**Customer Demographics**   

**Plot 1. Distribution of Purchase**
```{r message=FALSE, warning=FALSE}
demo %>%
  ggplot(aes(Purchase)) +
  geom_histogram(fill = "#014d64") +
  scale_x_continuous(limits = c(0, 11000000)) +
  theme_economist(base_size=14)+
  scale_fill_economist()+
  labs(x = "Purchase in Dollars", y = "Number of Customers", title = "Purchase Distribution")
```
[1]Glorious Christian, Black Friday Analysis https://www.kaggle.com/gloriousc/black-friday-analysis

**Plot 2. Purchase Distribution by Gender and City (Gloria)**
```{r message=FALSE, warning=FALSE}
ggthemr('flat dark')
demo %>%
  ggplot(aes(x = Purchase, fill = Gender)) +
  geom_histogram() +
  facet_wrap(~City_Category) +
  scale_x_continuous(limits = c(0, 3000000)) +
  labs(x = "Purchase in Dollars", y = "Number of Customers", title = "Purchase Distribution by Gender and City")+
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5))
```
**Plot 3. Purchase Distribution by Occupation (Vincent)**
```{r message=FALSE, warning=FALSE}
#Create new data frame grouping by occupation and marital status, and taking the mean purchase of each occupation and status.
ggthemr_reset()
demo %>%
  group_by(Occupation) %>%
  ggplot(aes(x = fct_infreq(Occupation), y = sum(Purchase), fill = Marital_Status)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_flip() +
  labs(x = "Occupation", y = "Purchase in Dollars", title = "Purchase Distribution by Occupation", fill="Marital Status") +
    scale_fill_discrete(labels=c("Unmarried", "Married"))
```
**Plot 4. Purchase Distribution by Age (Gloria)**
```{r message=FALSE, warning=FALSE}
ggthemr('sky')
demo %>%
  ggplot(aes(x = Age, y = Purchase)) +
  geom_boxplot() +
  labs(x = "Age", y = "Purchase in Dollars", title = "Purchase Distribution by Age")
```
**Plot 5. Purchase distribution by stay in city years (Zilei)**
```{r message=FALSE, warning=FALSE}
ggthemr('sky', layout = 'minimal')
demo %>%
  ggplot(aes(x = Stay_In_Current_City_Years, y = Purchase)) +
  geom_boxplot()+
  labs(x = "Stay in Current City Years", y = "Purchase", title = "Purchase Distribution by Stay Years")
```
In the above plot, we want to find the purchase distribution by stay in current city years. Because the stay in current city years is catergory and the purchase is continuous, we choose boxplot to visualize the relationship. We use different theme to make the plot more beautiful and add labels and titles to make it clearer. From this plot, we find there may be no difference among the purchase distribution of different years. But we want to have more confidence to draw such conclusion. So we use anova test in the next part to find whether it is significant. Based on the above plot, we draw the conclusion that "stay in current city years" may not influence the purchase and the store may not need to take this factor into consideration when making promotion decision.  

**Plot 6. The most popular product category**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggthemr('light', layout = 'clean')
BF_prt = BF %>%
  group_by(Product_ID) %>%
  mutate(sumpp = sum(Purchase), avgpp = mean(Purchase))
BF_prt = BF_prt[-c(3:8)]
BF_prt%>%
  ggplot(aes(x = Product_Category_1))+
  geom_bar(aes(x = fct_infreq(as.factor(Product_Category_1))))+
  labs(x ="Product Category" , y = "Purchase Amount", title = "Product Category 1 Distribution")
```
In the above visualization, we decided to choose Product Category 1 and Purchase variables in order to see which products were the highest selling. We decided to omit product 2 and 3 categories because many of the values are zero. Using a bar plot visualization is the most appropriate because it shows continuous data on the y-axis (purchase, using geom_bar) and discrete on the x-axis. We improved the plot by reordering the bars by highest to lowest and changing the theme colors.

We can easily see that the top 3 product 1 categories are 5, 1, and 8, by far. This visualization can help the business because they can direct more marketing dollars toward promoting the top 3 product categories. On the other hand, the company could use this plot to target certain products that it wants to sell more and direct marketing dollars towards those categories.

**4. Preliminary statistical analyses**  
**Do men spend more money on Black Friday than women?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Independent t-test
t.test(demo[demo$Gender == "M", ]$Purchase,
       demo[demo$Gender == "F", ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```
In the above code, we wanted to determine whether gender influences purchase. In table 1 and plot 2, we find that there are more men than women and that men have more purchases than women. Because the gender is category and the purchase is continuous, we choose t-test for figuring out whether there is a difference between male and female in purchase.  
Based on the evidence from this data set and the result from the t-test (p-value < 2.2e-16), we reject the null hypothesis and accept the alternative hypothesis: men have larger mean purchase than women. 

From the statistic analysis, we can give some recommendations to the store. (1) The company need to focus more on male customers because they have stronger purchase power than female and have more methods of promotion towards male customers to increase their revenue.

**Do unmarried customers spend more money on Black Friday than married?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Independent t-test
t.test(demo[demo$Marital_Status == 0, ]$Purchase,
       demo[demo$Marital_Status == 1, ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```
In this instance, we want to determine whether marital status influences purchase. In table 1 and plot 3, we find that there are more unmarried people than married, and that unmarried customers seem to make more purchases than married customers. Because maritial status is category and purchase is continuous, we use an independent t-test to test whether this difference is significant.  

Based on the evidence from this data set and the result from the t-test (p-value = 0.056), we draw the conclusion that in the 95% confidence interval, we cannot reject the null hypothesis. But in the 90% confidence interval, however, we can reject the null hypothesis and accept the alternative hypothesis: that unmarried customers have larger mean purchase than married customers.

Because the mean purchase is higher for unmarried people than married people, the company's marketing department could focus on targeting single people in order to increase purchases. However, it would be wise not to focus all marketing on single people, since the difference between the mean purchase amount for unmarried versus married is not stark.

**Is there an association between city and purchase?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Chi-Square test
# Create purchase categories
A1 = quantile(demo$Purchase, c(0.33,0.66))
A1
demo = demo %>% 
  mutate(purchase_category = cut(Purchase, breaks = c(-Inf, 304987.6, 826809.6, Inf),
  labels = c("Low", "Medium", "High")))

# Create two-way frequency table
table(demo$City_Category, demo$purchase_category)

# Apply Chi-Square test
chisq.test(table(demo$Age, demo$purchase_category))
```
In the above code, we want to determine whether city stay influences purchase. In Plot 2, we find differences in purchase distribution. Therefore, we want to use a Chi-Square test to see whether this finding is significant. 

Based on the evidence from this data set and the result from the Chi-Square test (p-value < 2.2e-16), we believe that the probability that there is no difference among the three cities is less than 5%, and the probability that there is a difference among the three cities is bigger than 95%. Therefore, we reject the null hypothesis, and accept the alternative hypothesis: that there is a difference in purchase distribution between the three cities. Therefore, from this data set, we draw the conclusion that there is an association between city and purchase. 

In terms of a business perspective, it would be wise for the company to analyze which cities are more profitable, so that it can appropriately direct marketing dollars towards either the more profitable or less profitable cities. 

**Conclusion**
From our preliminary data analysis, we were able to determine customer demographics that were associated with higher purchase amounts. In particular, we found that men spend more money on Black Friday than women, the purchase amounts are higher in different cities (City C has higher purchase amounts), unmarried people spend more than married (but only slightly more), certain occupation categories spend more than others, and interestingly, that there was not a huge difference in mean purchase by age (26-35 category spends slightly more). Also interestingly, there is almost no perceivable difference in mean purchase by years in city.

By using statistical tests and visualizations, we are able to easily pinpoint particular demographics on which the company should direct its marketing dollars. Further analysis will illuminate these important demographics.
