---
title: 'Marketing Analytics Project: Black Friday'
author: "Weicheng Zhang, Zilei Zhang, Maura Oray, Gloria Zheng"
output: word_document
---

**1. Introduction** 

Every year in America, there is one day where retail stores make the most money. In fact, stores made over $5 billion online from customers on this day in 2017. This date is always the day after Thanksgiving, and it's called "Black Friday." In our dataset, a particular company wants to gain a better understanding about customer purchasing behavior, because Black Friday is the most profitable day of the year. The company also aims to compare different products to see which ones customers are buying more, and also the demographics of the customers. Having this information could be useful for marketing so that they can target particular customer segments for particular products, and also affect company purchasing decisions for the next year.  

In this project, we aim to:
* Look at the distribution of customer demographics  
* Determine which variables are highly correlated with purchase  
* Identify which products are high-selling  

The dataset comes from a competition hosted by Analytics Vidhya.

Group Github url: https://github.com/MauraO/Marketing-Analytics-Project
```{r message=FALSE, warning=FALSE, include=FALSE, results='hide'}
#Load libraries
library(VIM)
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(devtools)
devtools::install_github('cttobin/ggthemr')
library(ggthemr)

BF <- read_csv("BlackFriday.csv")
```
**2. Data Description**

**a. Describe the conceptual measure types of the different variables**

User_ID (discrete, nominal), Product_ID (discrete, nominal), Gender (discrete, nominal), Age (discrete, ordinal), Occupation (discrete, nominal), City_Category (discrete, nominal), Stay_in_Current_City_Years (continuous, ratio), Marital_Status (discrete, nominal), Product_Category_1 (discrete, nominal), Product_Category_2 (discrete, nominal), Product_Category_3 (discrete, nominal), and Purchase (continuous, ratio). 

**b. Mention all the steps you took to clean the data**

**Step 1. Check if the data contains missing values**
```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
sum(complete.cases(BF))
aggr(BF,cex.axis = .4)
```
In this visualization, we wanted to see which variables had the most NA values. We see that the NA values were only concentrated in Product Category 2 and Product Category 3 (red indicates the location of NA values). Because these two variables are sub categories of Product Category 1 and our further analysis does not use them, we deleted them and made a new data set later.

**Step 2. Check the structure of the data**
```{r results='hide', message=FALSE, warning=FALSE}
summary(BF)
str(BF)
```
In this part, we wanted to see the type of variables for later use (group_by). We found that there were no factors, which leads us to the next step.

**Step 3. Type coercion**
```{r results='hide',message=FALSE, warning=FALSE}
BF[1:11] = lapply(BF[1:11], factor)
```
The majority of the variables were either integers or characters. We decided to change the variables into factors (all except purchase) because the majority of them are categorical, and coercing to factors makes for easier and clearer analysis (group_by).  

**Step 4. Tidy the data**
```{r results='hide',message=FALSE, warning=FALSE}
#Select demographic data
demo = BF %>% 
  select(-starts_with("Product")) %>%
  group_by(User_ID) %>%
  mutate(Purchase = sum(Purchase))
demo = demo[!duplicated(demo$User_ID), ]
```
Because we decided to focus more on customer analysis instead of the product, and that Product Category 2 and Product Category 3 had a lot of missing data, we deleted all of the product category columns to make a clearer customer data set. Moreover, we found that the Customer IDs were repeated many times which is redundant and we needed a purchase sum. Therefore, we used customers’ whole purchase amount to replace the individual product purchase and deleted the redundant customer IDs.  

**3. Summary statistics and Data Visualizations** 
**Customer Demographics**   

**Visualization 1. Distribution of Purchase**
```{r message=FALSE, warning=FALSE}
par(mfrow=c(2,2))
demo %>%
  ggplot(aes(Purchase)) +
  geom_histogram(fill = "#014d64") +
  scale_x_continuous(limits = c(0, 11000000)) +
  theme_economist(base_size=14)+
  scale_fill_economist()+
  labs(x = "Purchase in Dollars", y = "Number of Customers", title = "Purchase Distribution")
```
Here we selected the variable “Purchase.” We decided to use a histogram to view the distribution of customers and purchase because it is the best way to display continuous data.  
From the graph, it shows a downward slope between number of customers and purchase, which illustrates that a larger number of customers spend less than a smaller number of customers who spend more.

[1]Glorious Christian, Black Friday Analysis https://www.kaggle.com/gloriousc/black-friday-analysis  

**Visualization 2. Purchase Distribution by Gender and City**
```{r message=FALSE, warning=FALSE}
ggthemr('flat dark')
demo %>%
  ggplot(aes(x = Purchase, fill = Gender)) +
  geom_histogram() +
  facet_wrap(~City_Category) +
  scale_x_continuous(limits = c(0, 3000000)) +
  labs(x = "Purchase in Dollars", y = "Number of Customers", title = "Purchase Distribution by Gender and City")+
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5))
```
Here we selected the variables “Gender” and “City” in order to view the purchasing distribution of different genders in different cities. We used the histogram visualization because “Purchase” is continuous. We were interested in comparing the purchasing pattern in different cities and genders. We also improved the plot by adding facets which divides purchasing into the 3 cities. It shows that at the same price, there are more customers in city "B," and fewer customers in city “A.” Moreover, at the same price, men spend more than women.  

**Visualization 3. Purchase Distribution by Occupation**
```{r message=FALSE, warning=FALSE}
#Create new data frame grouping by occupation and marital status, and taking the mean purchase of each occupation and status.
ggthemr_reset()
demo %>%
  group_by(Occupation) %>%
  ggplot(aes(x = fct_infreq(Occupation), y = sum(Purchase), fill = Marital_Status)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_flip() +
  labs(x = "Occupation", y = "Purchase in Dollars", title = "Purchase Distribution by Occupation", fill="Marital Status") +
    scale_fill_discrete(labels=c("Unmarried", "Married"))
```
We were interested in determining which occupations spend more and whether married customers spend more than unmarried customers. Since marital status and occupation are factors (discrete) variables, and purchase is a continuous variable, we used a barplot to show total purchase for each occupation, split by matital status. Without looking at the marital status, this barplot shows the purchase distribution among 21 occupations and can answer questions such as "How much do customers from occupation x spend on Black Friday?" Taking into account marital status, the plot answers whether married or unmarried customers spend more. We improved the plot by flipping the coordinates and reordering the purchase by occupation in ascending order.
 Generally, unmarried customers spend more on Black Friday and this holds true for most occupations. Specifically, occupations 17, 1, 7, 0 and 4 spend the most money, over $2 trillion, out of which occupation 4 purchases the most. Interestingly, for occupation 10, married customers hardly spend money on Black Friday.  

**Visualization 4. Purchase Distribution by Age**
```{r message=FALSE, warning=FALSE}
ggthemr('sky')
demo %>%
  ggplot(aes(x = Age, y = Purchase)) +
  geom_boxplot() +
  labs(x = "Age", y = "Purchase in Dollars", title = "Purchase Distribution by Age")
```
Here we wanted to determine if there are similar purchasing patterns in different age ranges. A boxplot is the most appropriate visualization because age is discrete and purchase is continuous. We improved the plot by setting the theme as sky. The boxplot can answer the question if is there a difference in median purchase among the 7 groups. Customers aged from 18-55 have similar median purchases except for group 36-35, which has the highest median purchase and range. Customers who are under 17 or over 55, however seem to spend a little bit less.

**Visualization 5. Purchase Distribution by Stay in City Years**
```{r message=FALSE, warning=FALSE}
ggthemr('sky', layout = 'minimal')
demo %>%
  ggplot(aes(x = Stay_In_Current_City_Years, y = Purchase)) +
  geom_boxplot()+
  labs(x = "Stay in Current City Years", y = "Purchase", title = "Purchase Distribution by Stay Years")
```
In the above plot, we wanted to find the purchase distribution by stay in current city years. Because stay in current city years is discrete and purchase is continuous, we chose a boxplot visualization. We used a different theme to make the plot more visually appealing and added labels and titles to make it clearer. From this plot, we found that there may be no difference among the purchase distribution of different current city years stay. But we wanted to have more evidence to draw such conclusion. We used the ANOVA test in the next part to determine whether this difference is significant. Based on the above plot, we drew the conclusion that stay in current city years may not influence the purchase and the store may not need to take this factor into consideration when evaluating marketing decisions. 

**Visualization 6. Number of Items in Each Product Category**    

```{r}
# Number of Items VS Product Category 1
BF %>% group_by(Product_Category_1) %>% 
  summarise(num_items=n_distinct(Product_ID)) %>%
  ggplot(aes(x = reorder(factor(Product_Category_1), -num_items), y=num_items))+
  geom_bar(stat = 'identity')+
  labs(x ="Product Category 1" , y = "Number of Items", title = "Number of Items in Product Category 1")
  
```

The assumptions here were that the quantity of distinct products within a certain category lead to popularity of this category and cheaper products have more substitutes. To verify our aforementioned guess, we came to the question that how many products are contained in each category. So we plot the quantity of  distinct products on the vertical axis. The output showed that category  8  stored the most products, followed by category 5 and 1. Now we saw why categories 1, 5 and 8 were so popular!

**Visualization 7. The Number of Customers in Product Category**
```{r}
# Number of Customers VS Product Category 1
BF %>% group_by(Product_Category_1) %>% 
  summarise(num_customer=n_distinct(User_ID)) %>%
  ggplot(aes(x = reorder(factor(Product_Category_1), -num_customer), y=num_customer))+
  geom_bar(stat = 'identity')+
  labs(x ="Product Category 1" , y = "Number of Customers", title = "Number of Customers VS Product Category 1")
  
```
We were also interested in the customer distribution against product category, i.e.,  how many  distinct customers have ever bought from a specific category? The number of distinct customers were counted for each category and plotted as y variable. The plot showed that categories 1,5 and 8 were still the three categories that had the most visits.

Bases on the series of plots, our conclusion is that the 3 most popular product categories have the most product items and customers.  

**Visualization 8. The Most Profitable Product Category**  

```{r}
# Average Expenditure VS Product Category 1
BF %>% group_by(Product_Category_1) %>% 
  summarise(sum_purchase = sum(Purchase)) %>%
  ggplot(aes(x = reorder(factor(Product_Category_1), -sum_purchase), y=sum_purchase))+
  geom_bar(stat = 'identity')+
  labs(x ="Product Category 1" , y = "Sum Expenditure", title = "Total Expenditure VS Product Category 1")
  
```
In addition to the number of items and the number of customers of each category, the profitability, measured by sum purchase amount spent on each category, was another concern. So we sumed up the purchased amount for each category and plotted it against each category. The above bar plot showed that the top 3 categories that customers spent most money on total are still category 5,1 and 8. Hence, we came to the conclusion that categories 5,1 and 8，the most popular categories based on previous graph, are still the most profitable categories. 

**4. Preliminary statistical analyses**  

**Test 1. Do men spend more money on Black Friday than women?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Independent t-test
t.test(demo[demo$Gender == "M", ]$Purchase,
       demo[demo$Gender == "F", ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```
In the above code, we wanted to determine whether gender influences purchase. In table 1 and plot 2, we find that there are more men than women and that men make more purchases than women. Because gender is discrete and purchase is continuous, we chose to apply a t-test to determine whether there is a difference between men and women in purchase.  
Based on the evidence from this data set and the result from the t-test (p-value < 2.2e-16), we rejected the null hypothesis and accepted the alternative hypothesis: that men have a larger mean purchase amount than women.  
From the statistical analysis, we recommend that the store should focus more on men because they have higher purchasing power than women and for the store to have more methods of promotion towards men in order to increase their revenue.


**Test 2. Do unmarried customers spend more money on Black Friday than married?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Independent t-test
t.test(demo[demo$Marital_Status == 0, ]$Purchase,
       demo[demo$Marital_Status == 1, ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```
In this instance, we wanted to determine whether marital status influenced purchase. In table 1 and visualization 3, we found that there were more unmarried people than married, and that unmarried customers seem to make more purchases than married customers. Because marital status is categorical and purchase is continuous, we applied an independent t-test to test whether this difference was significant.  

Based on the evidence from this data set and the result from the t-test (p-value = 0.056), we drew the conclusion that in the 95% confidence interval, we cannot reject the null hypothesis. However, in the 90% confidence interval, we can reject the null hypothesis and accept the alternative hypothesis: that unmarried customers have larger mean purchase amounts than married customers.

Because the mean purchase amount was higher for unmarried people than married people, the company's marketing department could focus on targeting single people in order to increase purchase amounts. However, it would be wise not to focus all marketing dollars on single people, since the difference between the mean purchase amount for unmarried versus married is not stark.  

**Test 3. Is there an association between city and purchase?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Mutate a new column
demo = demo %>% 
  mutate(purchase_category = cut(Purchase, breaks = c(-Inf, 304987.6, 826809.6, Inf),
  labels = c("Low", "Medium", "High")))
table(demo$City_Category, demo$purchase_category)

# Apply Chi-Square test
chisq.test(table(demo$City_Category, demo$purchase_category))
```
On the above code, we aimed to see whether city stay influences purchase amounts. In visualization 2, we saw the differences in purchase distribution. Because city stay is a categorical variable, it was appropriate to use a Chi-Square test to determine whether there was significant differences in purchase between cities.  

Based on the evidence from this data set and the result from the Chi-Square test (p-value < 2.2e-16), we believe that the probability that there is no difference among the three cities is less than 5%, and that the probability that there is a difference among the three cities is greater than 95%. Therefore, we rejected the null hypothesis and accepted the alternative hypothesis: that there is a difference in purchase distribution between the three cities. Therefore, from this data set, we conclude that there is an association between city and purchase.  

From a business perspective, it would be wise for the company to analyze which cities are more profitable, so that it can appropriately direct marketing dollars towards either the more profitable or less profitable cities, depending on the marketing campaign.  

**Test 4. Is there an association between Age and purchase?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Chi-Square test
# Create purchase categories
demo = demo %>% 
  mutate(purchase_category = cut(Purchase, breaks = c(-Inf, 304987.6, 826809.6, Inf),
  labels = c("low", "medium", "high")))

# Create two-way frequency table
table(demo$Age, demo$purchase_category)

# Apply Chi-Square test
chisq.test(table(demo$Age, demo$purchase_category))
```
In visualization 4, there was a slight difference between average purchase amounts in the 7 age bins. Therefore, we decided to test if the difference was signficant. Since both age and purchase categories are discrete variables with more than 2 categories, we applied the Chi-Square test. This test would be able to determine if there was an association between age range and purchase levels. We added another variable, purchase_category, which would show purchase amounts as Low, Medium, or High. The result showed that the p-value for this test was 0 so we rejected the null hypothesis. Therefore, we concluded that age range did affect purchase levels on Black Friday. For example, in group 0-17, the higher the purchase level, the less amount of customers, while in group 26-35 the number of custmers increased with the price levels.  

**Test 5. Is there an association between staying years and purchase?**

```{r}
stay = lm(Purchase ~ Stay_In_Current_City_Years, data = demo)
anova(stay)
pairwise.t.test(demo$Purchase, demo$Stay_In_Current_City_Years, p.adjust.method = "bonferroni")
```

The boxplot of stay in current city years exhibited highly similar patterns for each subgroup. Therefore, we were interested in testing the relationship between stay in a certain city with Black Friday purchase amounts at the population level. Stay was a factor variable with 5 categories and purchase was a continuous variable, so we used analysis of variance (ANOVA), which is the extesion of the t-test and can be used to test whether there are significant differences in more than 2 population means (Ralph, 2010). However, since ANOVA does not tell how the mean of one population's group is different than the others, we still needed to use a pairwise t-test. A pairwise t-test enabled us to compare multiple groups' means with the correlation pairwisely. The p-value table showed that the differences in average purchase amounts were not significantly different regardless of how long the customer had been living there. We deduced this because the all the p-values were greater than 0.05. Therefore, we concluded that stay in current city years did not influence customers' purchasing behavior on Black Friday.  

**5. Regression Analyses**  

**a. The baseline regression model:**  
In part 4, we figured out gender, city and age are three significant factors, and marital status is a marginal significnat one. Therefore, we used the first three as the first baseline regressors and built our model.  

**b. Dependent and Independent variables**  
The dependent variable in our model is log(purchase), and independent variables are: gender, age, city category. Since all the demographic regressors are categorical variables, logarithm of purchase amount can better tell percentage change for different group of customers.
                           
**c. Data Aggregations**  
We used overall expenditure for each distinct User_ID. The cleaned data we did above which is data "demo". The data "demo" is data aggregation which we looked at overall expenditures for individual user. We used customers’ whole purchase amount to replace the individual product purchase.  

**d. Results of the Regression**  
```{r, echo=TRUE, message=FALSE, warning=FALSE}
lm_multiple_1=lm(log(Purchase)~Gender+Age+City_Category,data=demo)
summary(lm_multiple_1)
confint(lm_multiple_1)
```  
The output of baseline regression model shows that gender, age and city category have significant effect on purchase amount.
Male customers would purchase 28.3% more than female customers; customers aged from 26-35 would buy most on Black Friday, followed by cutomers in age range 36-45 and 45-50; customers from city B would purchase the most, followed by customers from city A and C. Due to the low R square coefficient, we focus more on the qualitative analysis rather than interpreting the coefficients of regressors.  

**e.Analyzing Controlling for omitted variable biases**  
The definition of omitted variable is that: (1) the omitted variavle is correlated with the included regressors and (2) the omitted variable is a determinent of the dependent variable. To verify the presence of omitted variable biases, we add one more unincluded variable one time to see if aforementioned effects would happen. The variable we add is used as the control, which can be defined as omitted variable if it significantly differs with 0 and and affect coefficient of other regressors.  
 
First, we added marital status as the control variable and re-do the regression 1.  
```{r}
lm_multiple_2=lm(log(Purchase)~Gender+Age+City_Category+Marital_Status,data=demo)
summary(lm_multiple_2)
```  

According to the result, we found that the marital status is not significant at all in this model. In other word, marital status is not a determinent of log(Purchase). Moreover, the R-square had no change, meaning that marital status can not explain the variance in dependent variable. Therefore, we can draw the conclusion that the marital status is not an omitted variable. 

Second, we added "stay_in_current_years" as the control variable and re-do the regression_1.
```{r}
lm_multiple_3=lm(log(Purchase)~Gender+Age+City_Category+Stay_In_Current_City_Years,data=demo)
summary(lm_multiple_3)
```
Again, although the values of coefficients of the regressors changed slightly, the variable "stay_in_current_years" is not significant in this mode. Moreover, the R-square hardly changed. Therefore, we can draw the conclusion that the marital status is not an omitted variable, for it had no influence on the other regressors and the dependent variable.  

Lastly, we added occupation as the third control variable, and re-do the regression_1. 
```{r}
lm_multiple_4=lm(log(Purchase)~Gender+Age+City_Category+Occupation,data=demo)
summary(lm_multiple_4)
```
Compared with result of regression 1 , we found that some occupations (3,6,7,13,16,17,19) are significant at 10% significance level. That is to say, occupation is correlated with purchase amount  to some degree. More importantly, customer aged from 18-25 and 46-50 are not marginally significant as before. Occupation is also correlated with age in some sense. Therefore, occupation is an omitted variable. This can be also approved by the increase in R square, from 0.1458 to 0.1512.  

**f.**  
Another data aggregation level we used to do regressions is user-product category level. The purcahse amount of products in the same category was aggregated for each customer. The purchase amount in user level is customer specific while at user-category level the purchase amount is customer and product categoory specific. At this data aggregation level we can answer questions from a more specific perspective: given a certain product category, how does  demographic factors affect the purchase amount; which category does man/ customer from city A/ customer aged from 18-25 purchase more? Or given a category which group of customers bought more.


```{r}
BF_cat = BF %>%
  group_by(Product_Category_1, User_ID) %>%
  mutate(sumpp = sum(Purchase))
BF_cat = BF_cat[-c(2)]
BF_cat = BF_cat[-c(9:11)]
BF_cat = BF_cat[!duplicated(BF_cat), ]

re_f <- lm(log(sumpp)~Age*Product_Category_1+Gender*Product_Category_1+City_Category*Product_Category_1+Occupation*Product_Category_1, data = BF_cat)
summary(re_f)


```
Product category 5 is a popular category so we take it as an example. The result shows that customers who are older than 55 purchased significant less than customers in other age ranges, and customer who are from city C spent less on this category than their counterparts.

product category 9 is a really interesting product category. The output shows that all the demographic regressors are not significant. The interpretation is that no matter which city the customer are from, whether the customers are young or old or whether they are male or female, there is no significant difference on purchase amount for this category. Therefore, we assumed that it may be a basic necessity of life for which all the customers have similar needs. 

**6. Segmentation:**  
**a.**  
The final model we used for segmentation is regressing log(Purchase) on age, gender, city and occupation. In other words, we can segment customers into male customers and female customers; or we can segment customers based on their age ranges; or we can segment customers based on their resident cities; or we can segment customers based on their occupations.  
Qualitatively, male segment purchased significanly more than female segment; customers in age range 26-35 is the segment that spent most on Black Friday, follwed by customer aged from 36-45 and the segment composed of all the rest customers; customers from city B was the top 1 segment in terms of purchase amount, followed by customers from city A and C; top 3 segments in terms of occupation were occupation 19, 3 and 16.


**b.**  
```{r message=FALSE, warning=FALSE}
# Expand factors into a set of dummy variables
data = model.matrix(~Gender+Age+Occupation+City_Category+Marital_Status+Stay_In_Current_City_Years,data = demo)
# Scale the data
scaled_data = scale(data,scale=FALSE)
k.max <- 15
wss <- sapply(1:k.max, 
              function(k){kmeans(scaled_data, k, nstart=50,iter.max = 20 )$tot.withinss})
# Plot out scree plot, determine the number of clusters
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
seg.summ <- function(data, groups) {
  aggregate(data, list(groups), function(x) mean(as.numeric(x))) 
}  
set.seed(96743) 
# Perform cluster analysis usinng Kmeans()
seg.k4 <- kmeans(scaled_data, centers=4)
seg.summ(data[,-1], seg.k4$cluster)
# Comparing groups on 1 variable
boxplot(demo$Purchase ~ seg.k4$cluster, 
        xlab="Purchase", ylab="Segment", horizontal=TRUE)
# Visualizing the overall clusters
library(cluster)
clusplot(demo, seg.k4$cluster, color=TRUE, shade=TRUE, 
         labels=4, lines=0, main="K-means clust")
```
Above is a plot of sum of squared distance for k in range 1-15. The plot look like an arm and its 'elbow' gives the optimal number of clusters, which in this dataset is 4. The 4 segments differ in age, occupation, city category, marital status and stay years in current city.  
* Cluster 1: Customers in this group who were exclusively from city C had the lowest purchase amount on Black Friday. Most customers are male (71.69%), aged from 18 to 45 (64%) and resident in current city for less than 1 year (64%). They were almost evenly distributed in terms of marital status.  
* Cluster 2: Customers in this group had the second lowest purchase amount on Black Friday. The majority of them are male (71.35%) and unmarried (58%), aged from 18 to 45 (77%). All customers had stayed in current city for 2 years and most come from city C (54.15%), followed by city B (29.87%) and city A (15.98%).  
* Cluster 3: Customers in this group purchased second most on Black Friday. Most customers are male (72.74%) and married (80.85%). And they are young, i.e., two thirds were aged from 26-35. Roughly half of total customers were from city A and most (72) had stayed in current city over 3 years.  
* Cluster 4: Customers in this group bought most on Black Friday. Most customers are male (71%), aged from 18 to 45 (71%), from in city B (75%) and resident in current city for less than 1 year (88%). They were almost evenly distributed in terms of marital status.  

**Regression after Segmentation**  
```{r}
demo$Cluster = seg.k4$cluster

cluster_multiple=lm(log(Purchase)~Gender*Cluster+Age*Cluster+Occupation*Cluster+Stay_In_Current_City_Years*Cluster+Marital_Status*Cluster,data=demo)
summary(cluster_multiple)
```

In cluster 1 where customers exclusively come from city C, male customers spent 23.3% more than female customers. Customers aged from 26-35 bought most, 31% more than customers aged under 17, followed by those who are in age range 46-50 and 51-55 (at 10% significance level). Those who have been doing occupation 6 and 7 purchased significantly less than occupation 0. _Marital status has no effect on purchase!_  

In cluster 2 where all customers were new in current city, male customers spent 32.3% more than female customers. There is no difference in purchase among various age ranges. Only Occupation 13 spent significantly less (53%) than occupation 0. It is noticible that, customers from city B purchased marginally (14.5% ) less and customers from city C spent significantly (72.4%) less than customers from city A. Matiral Status was not correlated with purchase on Black Friday.  

In cluster 3, male customer spent 39% more than female customers. Age, marital status and stay years in current city had no effect on purchase amount on Black Friday. Occupation 19 had the highest purcahse, followed by occupation 3. Customers from city B spent most followed by those who came from city A and C.  

Male customers in this clustr bought more than female customers. Age, marital status, and stay years in current years (at 5% significance level) did not affect purchase. Customers from city B spent more than customers from city A.  
In general, marital status is not correlated with purchase amount on Black Friday. Like wise, age has no effect on total spending except cluster 1.  

**Conclusion**

From our preliminary data analysis, we were able to determine customer demographics that were associated with higher purchase amounts. In particular, we found that men spend more money on Black Friday than women, the purchase amounts are higher in different cities (City C has higher purchase amounts), unmarried people spend more than married (but only slightly more), certain occupation categories spend more than others, and interestingly, that there was not a huge difference in mean purchase by age (26-35 category spends slightly more). Also interestingly, there is almost no perceivable difference in mean purchase by years in city. For the marital status, its effect on purchase is ambiguous.  

Analysis was further by regressions. Based on previous takeaways, we built a baseline model, which regressed log(purchase) on gender, age and city category. Different with qualitative analysis of statistical tests, regression model output quantitative information about how customer demographics affect purchase amount on Black Friday. However, regressions often suffer from omitted variables biases that have an impact on dependent variable but are not included. To control omitted variable, the rest variables (stay years, marital status and occupation) were added to the baseline regression one by one, holding base regressors  constant each time. Finally, we found that occupation is an omitted variable because some occupations(3,6,7,13,16,17,19) are  statistically significant. The extended regression showed that male customers would purchase significantly more than female customers; customers aged from 26-35 would buy most on Black Friday and customers from city B would purchase the most.  

Customer segmentation provided more insight into customers by grouping customers based on similar characteristics or purchasing behavior, which can be classified as either a priori or post-hoc. The result of a priori segmentation was based on customer demographics characteristics. While post-hoc segmentation segmented customers into 4 clusters. By re-doing regressions for each cluster, we found that marital status is not correlated with purchase amount on Black Friday, which matches the result of regression part; age has no effect on total spending except cluster 1.  

After identifying the demographic factors  that have significant effect on purchase amount and distinct purchasing behaviors of different customer segments, there are some managerial takeaways from the data pattern and model results. First, the manager could put their marketing focus on customers who are male, aged from 26-35 and from city B. To put it more specifically, the group of customers characterized by cluster 4 are the target customers. Therefore, this firm could send more emails informing target customers of the recommendation products and promotion information.
