---
title: 'Marketing Analytics Project: Black Friday'
author: "Weicheng Zhang, Zilei Zhang, Maura Oray, Gloria Zheng"
output:
word_document: default
html_document:
---

**1. Introduction** 
**Objective:**
* Look at the distribution of customer demographics
* Determine which variables are highly correlated with purchase
* Identify which products are high-selling and the characteristics of their customers

The dataset comes from a competition hosted by Analytics Vidhya.

Group Github url: https://github.com/MauraO/Marketing-Analytics-Project
```{r Echo =TRUE,message=FALSE, warning=FALSE}
#Load libraries
library(VIM)
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(devtools)
devtools::install_github('cttobin/ggthemr')
library(ggthemr)

BF <- read_csv("BlackFriday.csv")
```
**2. Data Description**
**a. Describe the conceptual measure types of the different variables**

User_ID (discrete, nominal), Product_ID (discrete, nominal), Gender (discrete, nominal), Age (discrete, ordinal), Occupation (discrete, nominal), City_Category (discrete, nominal), Stay_in_Current_City_Years (continuous, ratio), Marital_Status (discrete, nominal), Product_Category_1 (discrete, nominal), Product_Category_2 (discrete, nominal), Product_Category_3 (discrete, nominal), and Purchase (continuous, ratio). 

**b. Mention all the steps you took to clean the data**
**Step 1. Check if the data contains missing values**
```{r Echo =TRUE,message=FALSE, warning=FALSE}
sum(complete.cases(BF))
aggr(BF,cex.axis = .4) #164,278 NA values
BF$Product_Category_2[is.na(BF$Product_Category_2)]<- 0 #0 - category_2 is missing
BF$Product_Category_3[is.na(BF$Product_Category_3)]<- 0 #0 - category_3 is missing
```
In the above visualization, we wanted to see which variables had the most NA values. We saw that the NA values were only concentrated in Product Category 2 and Product Category 3 (red indicates the location of NA values). We thus decided to fill the NA values with "0" as this indicates that there were no sub categories in those rows.

**Step 2. Check the structure of the data**
```{r Echo =TRUE, message=FALSE, warning=FALSE}
summary(BF)
str(BF)
```
**Step 3. Type coercion**
```{r Echo =TRUE,message=FALSE, warning=FALSE}
BF[1:11] = lapply(BF[1:11], factor)
```
The majority of the variables were either integers or characters. We decided to coerce the variables into factors (all except purchase) because the majority of them are categorical, and coercing to factors makes for easier and clearer analysis.

**Step 4. Tidy the data**
```{r Echo =TRUE,message=FALSE, warning=FALSE}
#Select demographic data
demo = BF %>% 
  select(-starts_with("Product")) %>%
  group_by(User_ID) %>%
  mutate(Purchase = sum(Purchase))

#5891 unique customers

demo = demo[!duplicated(demo$User_ID), ]
```
**3. Summary statistics and Data Visualizations** 
**Customer Demographics**   
**Table 1. Distribution of Gender and Marital Status**
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Two-way frequency table showing different combination of customers using gender and marital status  
customer = table(demo$Gender, demo$Marital_Status)
colnames(customer) = c("Unmarried", "Married")
rownames(customer) = c("Female", "Male")

addmargins(customer)
addmargins(prop.table(customer))
```
The two-way frequency table shows 4 combinations of customers. There are 5891 distinct customers in total, and 2470 of them are unmarried males; 1755 are married males; 947 are unmarried females; 719 are married females. 

In this two-way frequency table, we can deduce that there are more unmarried customers (58%) than married (42%) and more male customers (72%) than female (28%). Therefore, in the following analyses regarding gender and marital status, we will focus on customers' average purchase amount instead of the sum purchase, because the sum purchase will easily be influenced by the number of people, which may cause a biased conclusion. 

**Plot 1. Distribution of Purchase (Gloria)**
```{r echo=TRUE, message=FALSE, warning=FALSE}
demo %>%
  ggplot(aes(Purchase)) +
  geom_histogram(fill = "#014d64") +
  scale_x_continuous(limits = c(0, 11000000)) +
  theme_economist(base_size=14)+
  scale_fill_economist()+
  labs(x = "Purchase in Dollars", y = "Number of Customers", title = "Purchase Distribution")
```
**Plot 2. Purchase Distribution by Gender and City (Gloria)**
```{r message=FALSE, warning=FALSE}
ggthemr('flat dark')
demo %>%
  ggplot(aes(x = Purchase, fill = Gender)) +
  geom_histogram() +
  facet_wrap(~City_Category) +
  scale_x_continuous(limits = c(0, 3000000)) +
  labs(x = "Purchase in Dollars", y = "Number of Customers", title = "Purchase Distribution by Gender and City")+
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5))
```
**Plot 3. Purchase Distribution by Occupation (Vincent)**
```{r message=FALSE, warning=FALSE}
#Create new data frame grouping by occupation and marital status, and taking the mean purchase of each occupation and status.
ggthemr_reset()
demo %>%
  group_by(Occupation) %>%
  ggplot(aes(x = fct_infreq(Occupation), y = sum(Purchase), fill = Marital_Status)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_flip() +
  labs(x = "Occupation", y = "Purchase in Dollars", title = "Purchase Distribution by Occupation", fill="Marital Status") +
    scale_fill_discrete(labels=c("Unmarried", "Married"))
```
**Plot 4. Purchase Distribution by Age (Gloria)**
```{r message=FALSE, warning=FALSE}
ggthemr('sky')
demo %>%
  ggplot(aes(x = Age, y = Purchase)) +
  geom_boxplot() +
  labs(x = "Age", y = "Purchase in Dollars", title = "Purchase Distribution by Age")
```
**Plot 5. Purchase distribution by stay in city years (Zilei)**
```{r message=FALSE, warning=FALSE}
ggthemr('sky', layout = 'minimal')
demo %>%
  ggplot(aes(x = Stay_In_Current_City_Years, y = Purchase)) +
  geom_boxplot()+
  labs(x = "Stay in Current City Years", y = "Purchase", title = "Purchase Distribution by Stay Years")
```
**Plot 6. The most popular product category**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggthemr('light', layout = 'clean')
BF_prt = BF %>%
  group_by(Product_ID) %>%
  mutate(sumpp = sum(Purchase), avgpp = mean(Purchase))
BF_prt = BF_prt[-c(3:8)]
BF_prt%>%
  ggplot(aes(x = Product_Category_1))+
  geom_bar(aes(x = fct_infreq(as.factor(Product_Category_1))))+
  labs(x ="Product Category" , y = "Purchase Amount", title = "Product Category 1 Distribution")
```
In the above visualization, we decided to choose Product Category 1 and Purchase variables in order to see which products were the highest selling. We decided to omit product 2 and 3 categories because many of the values are zero. Using a bar plot visualization is the most appropriate because it shows continuous data on the y-axis (purchase, using geom_bar) and discrete on the x-axis. We improved the plot by reordering the bars by highest to lowest and changing the theme colors.

We can easily see that the top 3 product 1 categories are 5, 1, and 8, by far. This visualization can help the business because they can direct more marketing dollars toward promoting the top 3 product categories. On the other hand, the company could use this plot to target certain products that it wants to sell more and direct marketing dollars towards those categories.

**4. Preliminary statistical analyses**  
**Do men spend more money on Black Friday than women?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Independent t-test
t.test(demo[demo$Gender == "M", ]$Purchase,
       demo[demo$Gender == "F", ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```
In the above code, we wanted to determine whether gender influences purchase. In table 1 and plot 2, we find that there are more men than women and that men have more purchases than women. Therefore, we want to use an independent t-test to test whether this difference is signficant.  

Based on the evidence from this data set and the result from the t-test (p-value < 2.2e-16), we think the probability that men have same mean purchase as women is less than 5%, and the probability that men have larger mean purchase than women is bigger than 95%. The p-value is very low, therefore, we reject the null hypothesis: that men have the same mean purchase as women, and accept the alternative hypothesis: men do not have the same mean purchase as women, and in fact have larger mean purchase. Moreover, from the sample, we can get the estimated mean purchase of men and women: 911963.2 and 699054.0.

**Do unmarried customers spend more money on Black Friday than married?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Independent t-test
t.test(demo[demo$Marital_Status == 0, ]$Purchase,
       demo[demo$Marital_Status == 1, ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```
In this instance, we want to determine whether marital status influences purchase. In table 1 and plot 3, we find that there are more unmarried people than married, and that unmarried customers seem to make more purchases than married customers. Therefore, we want to use an independent t-test to test whether this difference is significant.  

Based on the evidence from this data set and the result from the t-test (p-value = 0.056), we think the probability that unmarried customers have same mean purchase as married customers is bigger than 5% but less than 10%, and that the probability that unmarried customers have larger mean purchase than married customers is less than 95% but bigger than 90%. Therefore, in the 95% confidence interval, we cannot reject the null hypothesis: that unmarried customers have the same mean purchase as married customers. In the 90% confidence interval, however, we can reject the null hypothesis and accept the alternative hypothesis: that unmarried customers have larger mean purchase than married customers.

Because the mean purchase is higher for unmarried people than married people, the company's marketing department could focus on targeting single people in order to increase purchases. However, it would be wise not to focus all marketing on single people, since the difference between the mean purchase amount for unmarried versus married is not stark.

**Is there an association between city and purchase?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Chi-Square test
# Create purchase categories
A1 = quantile(demo$Purchase, c(0.33,0.66))
A1
demo = demo %>% 
  mutate(purchase_category = cut(Purchase, breaks = c(-Inf, 304987.6, 826809.6, Inf),
  labels = c("City A", "City B", "City C")))

# Create two-way frequency table
table(demo$City_Category, demo$purchase_category)

# Apply Chi-Square test
chisq.test(table(demo$Age, demo$purchase_category))
```
In the above code, we want to determine whether city stay influences purchase. In Plot 2, we find differences in purchase distribution. Therefore, we want to use a Chi-Square test to see whether this finding is significant. 

Based on the evidence from this data set and the result from the Chi-Square test (p-value < 2.2e-16), we believe that the probability that there is no difference among the three cities is less than 5%, and the probability that there is a difference among the three cities is bigger than 95%. Therefore, we reject the null hypothesis, and accept the alternative hypothesis: that there is a difference in purchase distribution between the three cities. Therefore, from this data set, we draw the conclusion that there is an association between city and purchase. 

In terms of a business perspective, it would be wise for the company to analyze which cities are more profitable, so that it can appropriately direct marketing dollars towards either the more profitable or less profitable cities. 

**Conclusion**
From our preliminary data analysis, we were able to determine customer demographics that were associated with higher purchase amounts. In particular, we found that men spend more money on Black Friday than women, the purchase amounts are higher in different cities (City C has higher purchase amounts), unmarried people spend more than married (but only slightly more), certain occupation categories spend more than others, and interestingly, that there was not a huge difference in mean purchase by age (26-35 category spends slightly more). Also interestingly, there is almost no perceivable difference in mean purchase by years in city.

By using statistical tests and visualizations, we are able to easily pinpoint particular demographics on which the company should direct its marketing dollars. Further analysis will illuminate these important demographics.