---
title: "Marketing Anlytics Project"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Install and load all the packages we need
library('VIM')
library(readr)
library(tidyverse)
library(dplyr)
```

### 1. Introduction 

#### Objective:
* Look at the distribution of customers demographcis
* Find how is purchase differed w.r.t. demographcis factors and test it
* what kind of products are sold well and who is buying them (to be decided)

### 2. Data Description

#### a. Describe the ???conceptual??? measure types of the different variables
```{r}
BF =  read_csv("BlackFriday.csv")
```

* User_ID : discrete
* Product_ID : discrete
* Gender : discrete
* Age : discrete
* Occupation : discrete
* City_Category : discrete
* Stay_in_Current_City_Years : continuous
* Marital_Status : discrete
* Product_Category_1 : discrete
* Product_Category_2 : discrete
* Product_Category_3 : discrete
* and finally Purchase :continuous

#### b. Mention all the steps you took to clean the data

##### Step 1. Check if the data contains missing values
```{r}
sum(complete.cases(BF))
aggr(BF,cex.axis = .4)
```

##### Step 2. Check the structure of the data
```{r message=FALSE, warning=FALSE}
summary(BF)
str(BF)
```

#### Step 3. Type coercion
```{r}
BF[1:11] = lapply(BF[1:11], factor)
```

#### Step 4. Tidy the data
```{r}
# Demographics data
demo = BF %>% 
  select(-starts_with("Product")) %>%
  group_by(User_ID) %>%
  mutate(Purchase = sum(Purchase))

demo = demo[!duplicated(demo$User_ID), ]
```


### 3. Summary statistics and Data Visualizations

```{r}
# Two way frequency table showing different combination of customers using gender and marital status
customer = table(demo$Gender, demo$Marital_Status)
colnames(customer) = c("Unmarried", "Married")
rownames(customer) = c("Female", "Male")
customer
```
The two way frequency table shows 4 combinations of customers. There are 5891 distinct customers in total, and 2470 of them are unmarried male; 1755 are married male; 947 are unmarried female; 719 are married female. The pattern is that more unmarried customers (58%) than the married (42%) and more male customers (72%) than the female (28%).


```{r message=FALSE, warning=FALSE}
# Purchase disttribution by gender and city
demo %>%
  ggplot(aes(x = Purchase, fill = Gender)) +
  geom_histogram() +
  facet_grid(City_Category~.) +
  scale_x_continuous(limits = c(0, 3000000)) +
  theme_minimal() +
  labs(x = "Purchase in dollar", y = "Number of customers", title = "Purchase disttribution by gender and city")
```

```{r}
# Purchase disttribution by age
demo %>%
  ggplot(aes(x = Age, y = Purchase)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Age", y = "Purchase in dollar", title = "Purchase disttribution by age")
```

```{r}
# Purchase disttribution by occupation
demo %>%
  ggplot(aes(x = fct_infreq(Occupation), y = Purchase)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_flip() +
  labs(x = "Occupation", y = "Purchase in dollar", title = "Purchase disttribution by occupation")
```

### 4 Preliminary statistical analyses

```{r}
## Independent t test
# Does men spend more money on Black Friday than women?
t.test(demo[demo$Gender == "M", ]$Purchase,
       demo[demo$Gender == "F", ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```

```{r}
## Independent t test
# Does unmarried customers spend more money on Black Friday than the married?
t.test(demo[demo$Marital_Status == 0, ]$Purchase,
       demo[demo$Marital_Status == 1, ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```

```{r}
### Chi square test
## Is there an association between age and purchase?

# Creat purchase category
demo = demo %>% 
  mutate(purchase_category = cut(Purchase, breaks = c(-Inf, 234914, 512612, 1099005, Inf),
                                        labels = c("low", "middle", "high", "very high")))

# Creat two way frequency table
table(demo$Age, demo$purchase_category)

# Apply chi square test
chisq.test(table(demo$Age, demo$purchase_category))

```

