---
title: 'Marketing Analytics Project: Black Friday'
author: "Weicheng Zhang, Zilei Zhang, Maura Oray, Gloria Zheng"
output: word_document
---

**1. Introduction** 

Every year in America, there is one day where retail stores make the most money. In fact, stores made over $5 billion online from customers on this day in 2017[^1]. This date is always the day after Thanksgiving, and it's called "Black Friday." In our dataset, a particular company wants to gain a better understanding about customer purchasing behavior, because Black Friday is the most profitable day of the year. The company also aims to compare different products to see which ones customers are buying more, and also the demographics of the customers. Having this information could be useful for marketing so that they can target particular customer segments for particular products, and also affect company purchasing decisions for the next year. 

In this project, we aim to:
* Look at the distribution of customer demographics
* Determine which variables are highly correlated with purchase
* Identify which products are high-selling

The dataset comes from a competition hosted by Analytics Vidhya.

Group Github url: https://github.com/MauraO/Marketing-Analytics-Project

```{r message=FALSE, warning=FALSE, include=FALSE, results='hide'}
#Load libraries
library(VIM)
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(devtools)
devtools::install_github('cttobin/ggthemr')
library(ggthemr)

setwd('~/Desktop/BUS-256/Marketing Analytics Project')
BF <- read_csv("BlackFriday.csv")
```
**2. Data Description**

**a. Describe the conceptual measure types of the different variables**

User_ID (discrete, nominal), Product_ID (discrete, nominal), Gender (discrete, nominal), Age (discrete, ordinal), Occupation (discrete, nominal), City_Category (discrete, nominal), Stay_in_Current_City_Years (continuous, ratio), Marital_Status (discrete, nominal), Product_Category_1 (discrete, nominal), Product_Category_2 (discrete, nominal), Product_Category_3 (discrete, nominal), and Purchase (continuous, ratio). 

**b. Mention all the steps you took to clean the data**

**Step 1. Check if the data contains missing values**
```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
sum(complete.cases(BF))
aggr(BF,cex.axis = .4)
```

In the visualization in the appendix, we wanted to see which variables had the most NA values. We see that the NA values were only concentrated in Product Category 2 and Product Category 3 (red indicates the location of NA values). Because these two variables are sub categories of Product Category 1 and our further analysis does not use them, we deleted them and made a new data set later.

**Step 2. Check the structure of the data**
```{r results='hide', message=FALSE, warning=FALSE}
summary(BF)
str(BF)
```

In this part, we wanted to see the type of variables for later use (group_by). We found that there were no factors, which leads us to the next step.

**Step 3. Type coercion**
```{r results='hide',message=FALSE, warning=FALSE}
BF[1:11] = lapply(BF[1:11], factor)
```

The majority of the variables were either integers or characters. We decided to change the variables into factors (all except purchase) because the majority of them are categorical, and coercing to factors makes for easier and clearer analysis (group_by).

**Step 4. Tidy the data**
```{r results='hide',message=FALSE, warning=FALSE}
#Select demographic data
demo = BF %>% 
  select(-starts_with("Product")) %>%
  group_by(User_ID) %>%
  mutate(Purchase = sum(Purchase))
demo = demo[!duplicated(demo$User_ID), ]
```

Because we decided to focus more on customer analysis instead of the product, and that Product Category 2 and Product Category 3 had a lot of missing data, we deleted all of the product category columns to make a clearer customer data set. Moreover, we found that the Customer IDs were repeated many times which is redundant and we needed a purchase sum. Therefore, we used customers’ whole purchase amount to replace the individual product purchase and deleted the redundant customer IDs. 

**3. Summary statistics and Data Visualizations**
**Customer Demographics**   

**Visualization 1. Distribution of Purchase**
```{r message=FALSE, warning=FALSE}
par(mfrow=c(2,2))
demo %>%
  ggplot(aes(Purchase)) +
  geom_histogram(fill = "#014d64") +
  scale_x_continuous(limits = c(0, 11000000)) +
  theme_economist(base_size=14)+
  scale_fill_economist()+
  labs(x = "Purchase in Dollars", y = "Number of Customers", title = "Purchase Distribution")
```

Here we selected the variable “Purchase.” We decided to use a histogram to view the distribution of customers and purchase because it is the best way to display continuous data. 

From the graph, it shows a downward slope between number of customers and purchase, which illustrates that a larger number of customers spend less than a smaller number of customers who spend more.

[1]Glorious Christian, Black Friday Analysis https://www.kaggle.com/gloriousc/black-friday-analysis

**Visualization 2. Purchase Distribution by Gender and City**
```{r message=FALSE, warning=FALSE}
ggthemr('flat dark')
demo %>%
  ggplot(aes(x = Purchase, fill = Gender)) +
  geom_histogram() +
  facet_wrap(~City_Category) +
  scale_x_continuous(limits = c(0, 3000000)) +
  labs(x = "Purchase in Dollars", y = "Number of Customers", title = "Purchase Distribution by Gender and City")+
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5))
```

Here we selected the variables “Gender” and “City” in order to view the purchasing distribution of different genders in different cities. We used the histogram visualization because “Purchase” is continuous. We were interested in comparing the purchasing pattern in different cities and genders. We also improved the plot by adding facets which divides purchasing into the 3 cities. It shows that at the same price, there are more customers in city "B," and fewer customers in city “A.” Moreover, at the same price, men spend more than women.

**Visualization 3. Purchase Distribution by Occupation**
```{r message=FALSE, warning=FALSE}
#Create new data frame grouping by occupation and marital status, and taking the mean purchase of each occupation and status.
ggthemr_reset()
demo %>%
  group_by(Occupation) %>%
  ggplot(aes(x = fct_infreq(Occupation), y = sum(Purchase), fill = Marital_Status)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_flip() +
  labs(x = "Occupation", y = "Purchase in Dollars", title = "Purchase Distribution by Occupation", fill="Marital Status") +
    scale_fill_discrete(labels=c("Unmarried", "Married"))
```

We were interested in determining which occupations spend more and whether married customers spend more than unmarried customers. Since marital status and occupation are factors (discrete) variables, and purchase is a continuous variable, we used a barplot to show total purchase for each occupation, split by matital status. Without looking at the marital status, this barplot shows the purchase distribution among 21 occupations and can answer questions such as "How much do customers from occupation x spend on Black Friday?" Taking into account marital status, the plot answers whether married or unmarried customers spend more. We improved the plot by flipping the coordinates and reordering the purchase by occupation in ascending order.

Generally, unmarried customers spend more on Black Friday and this holds true for most occupations. Specifically, occupations 17, 1, 7, 0 and 4 spend the most money, over $2 trillion, out of which occupation 4 purchases the most. Interestingly, for occupation 10, married customers hardly spend money on Black Friday. 

**Visualization 4. Purchase Distribution by Age**
```{r message=FALSE, warning=FALSE}
ggthemr('sky')
demo %>%
  ggplot(aes(x = Age, y = Purchase)) +
  geom_boxplot() +
  labs(x = "Age", y = "Purchase in Dollars", title = "Purchase Distribution by Age")
```

Here we wanted to determine if there are similar purchasing patterns in different age ranges. A boxplot is the most appropriate visualization because age is discrete and purchase is continuous. We improved the plot by setting the theme as sky. The boxplot can answer the question if is there a difference in median purchase among the 7 groups. Customers aged from 18-55 have similar median purchases except for group 36-35, which has the highest median purchase and range. Customers who are under 17 or over 55, however seem to spend a little bit less.

**Visualization 5. Purchase Distribution by Stay in City Years**
```{r message=FALSE, warning=FALSE}
ggthemr('sky', layout = 'minimal')
demo %>%
  ggplot(aes(x = Stay_In_Current_City_Years, y = Purchase)) +
  geom_boxplot()+
  labs(x = "Stay in Current City Years", y = "Purchase", title = "Purchase Distribution by Stay Years")
```

In the above plot, we wanted to find the purchase distribution by stay in current city years. Because stay in current city years is discrete and purchase is continuous, we chose a boxplot visualization. We used a different theme to make the plot more visually appealing and added labels and titles to make it clearer. From this plot, we found that there may be no difference among the purchase distribution of different current city years stay. But we wanted to have more evidence to draw such conclusion. We used the ANOVA test in the next part to determine whether this difference is significant. Based on the above plot, we drew the conclusion that stay in current city years may not influence the purchase and the store may not need to take this factor into consideration when evaluating marketing decisions. 

**Visualization 6. The Most Popular Product Category**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggthemr('light', layout = 'clean')
BF%>%
  ggplot(aes(x = Product_Category_1))+
  geom_bar(aes(x = fct_infreq(as.factor(Product_Category_1))))+
  labs(x ="Product Category" , y = "Purchase Frequency", title = "The Distribution of Product Category 1 Frequency")
```

In the above visualization, we decided to choose Product Category 1 and Purchase variables in order to see which products were the highest selling. We decided to omit Product 2 and 3 categories because many of the values were zero. Using a bar plot visualization is the most appropriate because it shows continuous data on the y-axis (purchase, using geom_bar) and discrete on the x-axis. We improved the plot by reordering the bars by highest to lowest and changing the theme colors.

We can easily see that the top 3 product 1 categories are 5, 1, and 8. This visualization can help the business direct more marketing dollars toward promoting the top 3 product categories. The company could use this plot to target certain products that it wants to sell more and direct marketing dollars towards those categories.

**Visualization 7. The Most Profitable Product Category**  
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggthemr('light', layout = 'clean')
BF_prt_1 = BF %>%
  group_by(Product_Category_1, User_ID) %>%
  mutate(sumpp = sum(Purchase))%>%
  select(Product_Category_1, avgpp)
BF_prt_1 = BF_prt_1[!duplicated(BF_prt_1$Product_Category_1), ]

BF_prt_1%>%
  ggplot(aes(x = Product_Category_1, y=avgpp))+
  geom_bar(stat = 'identity')+
  labs(x ="Product Category" , y = "Purchase Cost", title = "The Distribution of Product Category 1's Profitability")
```

In the above visualization, we decided to futher analyze whether the most popular categories (5,1, and 8) are also the most profitable categories. We calculated the average purchase amount in each category in Product Category 1. Because we wanted to see the comparison between different product categories, we decided to use a bar plot.  

After looking at the previous visualization, we initially thought that categories 5,1, and 8 were the most popular categories. However, when we plotted the distribution of average purchase amounts of Product Category 1, we found that the most popular categories were not the most profitable.  

**Visualization 8. Number of Items in Each Product Category**  
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggthemr('light', layout = 'clean')
BF_prt_2 = BF %>%
  select(Product_Category_1, Product_ID)
BF_prt_2 = BF_prt_2[!duplicated(BF_prt_2$Product_ID), ]
BF_prt_2 = count(BF_prt_2,Product_Category_1)

BF_prt_2%>%
  ggplot(aes(x = Product_Category_1, y=n))+
  geom_bar(stat = 'identity')+
  labs(x ="Product Category" , y = "Customer Number", title = "The Number of Items in Product Category 1")
```  

In the above visualization, we decided to examine why categories 5,1, and 8 were the most frequently purchased. We assumed that the number of items in these categories could lead to their popularity (for example, if the product has a developed and extensive product line, it may become more popular). Therefore, we created a plot to see how many items each product category contained. Furthermore, because we wanted to easily see the comparison between categories, a bar plot was the most appropriate choice.  

After examining this visualization, we could deduce a possible reason why categories 5,1, and 8 were the most popular categories. These categories contain more products than the other categories. If they contain more products, the customer's probability of purchasing and item from one of these three categories is much more likely.

**Visualization 9. Number of Customers in Product Category**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# 如果一个顾客买了好几次同一个category, 算几次？下面这个代码只算了一次
# for example, User ID 1000004, bought product category 1 for 3 times. (but not the same produ ct.)
ggthemr('light', layout = 'clean')
BF_prt_3 = BF %>%
  select(Product_Category_1, User_ID)
BF_prt_3 = BF_prt_3[!duplicated(BF_prt_3), ]
BF_prt_3 = count(BF_prt_3,Product_Category_1)


BF_prt_3%>%
  ggplot(aes(x = Product_Category_1, y=n))+
  geom_bar(stat = 'identity')+
  labs(x ="Product Category" , y = "Customer Number", title = "Number of Customers in Product Category 1")
```

Once again, it is clear that the number of customers are the highest in categories 5,1, and 1, which further cements our assumption that the most popular products are also the most profitable, and have the most items and number of customers. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
BF_prt_4 = BF %>%
  group_by(Product_Category_1) %>%
  mutate(sumpp = sum(Purchase))%>%
  select(Product_Category_1, sumpp)
BF_prt_4 = BF_prt_4[!duplicated(BF_prt_4$Product_Category_1), ]

category = inner_join(BF_prt_2, BF_prt_3, by = "Product_Category_1")
category = inner_join(category, BF_prt_4, by = "Product_Category_1")

names(category) <-c('Product_Category', 'Item_Number','Customer_Number', 'Purchase Amount')

## 我们还有其他aggregation的方法 但是sample size太小了 所以不做回归了
```
**4. Preliminary statistical analyses**  

**Test 1. Do men spend more money on Black Friday than women?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Independent t-test
t.test(demo[demo$Gender == "M", ]$Purchase,
       demo[demo$Gender == "F", ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```
In the above code, we wanted to determine whether gender influences purchase. In table 1 and plot 2, we find that there are more men than women and that men make more purchases than women. Because gender is discrete and purchase is continuous, we chose to apply a t-test to determine whether there is a difference between men and women in purchase.  
Based on the evidence from this data set and the result from the t-test (p-value < 2.2e-16), we rejected the null hypothesis and accepted the alternative hypothesis: that men have a larger mean purchase amount than women.  
From the statistical analysis, we recommend that the store should focus more on men because they have higher purchasing power than women and for the store to have more methods of promotion towards men in order to increase their revenue.

**Test 2. Do unmarried customers spend more money on Black Friday than married?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Independent t-test
t.test(demo[demo$Marital_Status == 0, ]$Purchase,
       demo[demo$Marital_Status == 1, ]$Purchase,
       paired = FALSE,
       alternative = "greater")
```
In this instance, we wanted to determine whether marital status influenced purchase. In table 1 and visualization 3, we found that there were more unmarried people than married, and that unmarried customers seem to make more purchases than married customers. Because marital status is categorical and purchase is continuous, we applied an independent t-test to test whether this difference was significant.  

Based on the evidence from this data set and the result from the t-test (p-value = 0.056), we drew the conclusion that in the 95% confidence interval, we cannot reject the null hypothesis. However, in the 90% confidence interval, we can reject the null hypothesis and accept the alternative hypothesis: that unmarried customers have larger mean purchase amounts than married customers.

Because the mean purchase amount was higher for unmarried people than married people, the company's marketing department could focus on targeting single people in order to increase purchase amounts. However, it would be wise not to focus all marketing dollars on single people, since the difference between the mean purchase amount for unmarried versus married is not stark.

**Test 3. Is there an association between city and purchase?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Mutate a new column
demo = demo %>% 
  mutate(purchase_category = cut(Purchase, breaks = c(-Inf, 304987.6, 826809.6, Inf),
  labels = c("Low", "Medium", "High")))
table(demo$City_Category, demo$purchase_category)

# Apply Chi-Square test
chisq.test(table(demo$Age, demo$purchase_category))
```
In the above code, we aimed to see whether city stay influences purchase amounts. In visualization 2, we saw the differences in purchase distribution. Because city stay is a categorical variable, it was appropriate to use a Chi-Square test to determine whether there was significant differences in purchase between cities.

Based on the evidence from this data set and the result from the Chi-Square test (p-value < 2.2e-16), we believe that the probability that there is no difference among the three cities is less than 5%, and that the probability that there is a difference among the three cities is greater than 95%. Therefore, we rejected the null hypothesis and accepted the alternative hypothesis: that there is a difference in purchase distribution between the three cities. Therefore, from this data set, we conclude that there is an association between city and purchase. 

From a business perspective, it would be wise for the company to analyze which cities are more profitable, so that it can appropriately direct marketing dollars towards either the more profitable or less profitable cities, depending on the marketing campaign. 

**Test 4. Is there an association between Age and purchase?**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Chi-Square test
# Create purchase categories
demo = demo %>% 
  mutate(purchase_category = cut(Purchase, breaks = c(-Inf, 304987.6, 826809.6, Inf),
  labels = c("low", "medium", "high")))

# Create two-way frequency table
table(demo$Age, demo$purchase_category)

# Apply Chi-Square test
chisq.test(table(demo$Age, demo$purchase_category))
```
In visualization 4, there was a slight difference between average purchase amounts in the 7 age bins. Therefore, we decided to test if the difference was signficant. Since both age and purchase categories are discrete variables with more than 2 categories, we applied the Chi-Square test. This test would be able to determine if there was an association between age range and purchase levels. We added another variable, purchase_category, which would show purchase amounts as Low, Medium, or High. The result showed that the p-value for this test was 0 so we rejected the null hypothesis. Therefore, we concluded that age range did affect purchase levels on Black Friday. For example, in group 0-17, the higher the purchase level, the less amount of customers, while in group 26-35 the number of custmers increased with the price levels. 

**Test 5. Is there an association between staying years and purchase?**
```{r}
stay = lm(Purchase ~ Stay_In_Current_City_Years, data = demo)
anova(stay)
pairwise.t.test(demo$Purchase, demo$Stay_In_Current_City_Years, p.adjust.method = "bonferroni")
```

The boxplot of stay in current city years exhibited highly similar patterns for each subgroup. Therefore, we were interested in testing the relationship between stay in a certain city with Black Friday purchase amounts at the population level. Stay was a factor variable with 5 categories and purchase was a continuous variable, so we used analysis of variance (ANOVA), which is the extesion of the t-test and can be used to test whether there are significant differences in more than 2 population means (Ralph, 2010). However, since ANOVA does not tell how the mean of one population's group is different than the others, we still needed to use a pairwise t-test. A pairwise t-test enabled us to compare multiple groups' means with the correlation pairwisely. The p-value table showed that the differences in average purchase amounts were not significantly different regardless of how long the customer had been living there. We deduced this because the all the p-values were greater than 0.05. Therefore, we concluded that stay in current city years did not influence customers' purchasing behavior on Black Friday.

**5. Regressions**

a. The baseline regression model: nonlinear regression model.We decided to use the nature logarithm of Y as an example of nonlinear regresion model. We wanted to examine the purchase gap between gender, age, occupation, city, the years of staying in the city,and the marital status. It is easier to compare purchase gaps when they are expressed in percentage terms.

b. Dependent variable Y: log(purchase) 
   Independent variable x: gender -dummy varible
                           age-dummy vairable
                           occupation-dummy variable
                           city category-dummy variable
                           stay_in_current_city_years-dummy variable
                           marital status-dummy variable
c. We used overall expenditure for each distinct User_ID. We used the cleaned data we did above which is data "demo". The data "demo"" is data aggregation which we looked at overall expenditures for individual user. We used customers’ whole purchase amount to replace the individual product purchase.

The below is the simple nonlinear regression model for each independent variable.
```{r, echo=TRUE, message=FALSE, warning=FALSE}                          
lm_gender=lm(log(Purchase)~Gender,data=demo)
summary(lm_gender)
```
### Explnation: We run a regression of log purchase on the dummy variable signifying the difference between male and female. We found when the individaul is male compared to when indinvidal is female, the individual purchase amount increase significanly by 27.49% at 99.99% significant level.

```{r, echo=TRUE, message=FALSE, warning=FALSE} 
lm_age=lm(log(Purchase)~Age,data=demo)
summary(lm_age)
```
### Explnation: We run a regression of log purchase on the dummy variable signifying the difference between different age range. We found when the individaul is 18-25 compared to when indinvidal is 0-17, the individual purchase amount increase significanly by 26.49% at 99.99% significant level.The individaul is 26-35 compared to when indinvidal is 0-17, the individual purchase amount increase significanly by 39.51% at 99.99% significant level.The individaul is 36-45 compared to when indinvidal is 0-17, the individual purchase amount increase significanly by 27.47% at 99.99% significant level.The individaul is 46-50 compared to when indinvidal is 0-17, the individual purchase amount increase significanly by 17.32% at 99% significant level.The individaul is 51-55 compared to when indinvidal is 0-17, the individual purchase amount increase significanly by 16.52% at 99% significant level.The individaul is 55+ compared to when indinvidal is 0-17, the individual purchase amount decrease significanly by 14.25% at 95% significant level.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
lm_occupation=lm(log(Purchase)~Occupation,data=demo)
summary(lm_occupation)
```
Explnation: We run a regression of log purchase on the dummy variable signifying the difference between different occupation. We found when the individaul is occupation1 compared to when indinvidal is occupation0, the individual purchase amount decrease significanly by 10.12% at 95% significant level.The individaul is occupation 9 compared to when indinvidal is occupation 0, the individual purchase amount decrease significanly by 38.20% at 99.99% significant level.The individaul is occupation 10 compared to when indinvidal is occupation 0, the individual purchase amount decrease significanly by 27.65% at 99.99% significant level.The individaul is occupation13 compared to when indinvidal is occupation 0, the individual purchase amount decrease significanly by 45.28% at 99.99% significant level.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
lm_citycategory=lm(log(Purchase)~City_Category,data=demo)
summary(lm_citycategory)
```
### Explnation: We run a regression of log purchase on the dummy variable signifying the difference between different city categories. We found when the individaul is in city B compared to when indinvidal is in city A, the individual purchase amount increase significanly by 10.34% at 99.9% significant level.The individaul is in city C compared to when indinvidal is in city A, the individual purchase amount decrease significanly by 62.09% at 99.99% significant level.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
lm_stayincityyears=lm(log(Purchase)~Stay_In_Current_City_Years,data=demo)
summary(lm_stayincityyears)
```

### Explnation: We run a regression of log purchase on the dummy variable signifying the difference between different years staying in current city. We found in 95% significant level, stay_in_current_city_years are not significantly influence the percentage change in purchae amount.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
lm_marital=lm(log(Purchase)~Marital_Status,data=demo)
summary(lm_marital)
```
### Explnation: We run a regression of log purchase on the dummy variable signifying the difference between married and unmarried. We found when the individaul is married compared to when indinvidal is unmarried, the individual purchase amount decrease significanly by 4.70% at 95% significant level.

d. Then we do the multiple nonlinear regresion model.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
lm_multiple=lm(log(Purchase)~Gender+Age+Occupation+City_Category+Stay_In_Current_City_Years+Marital_Status
+Gender,data=demo)
summary(lm_multiple)
```
### Explanation: We run a regression of log purchase on all the independent variables together. We found at 99.99% significance level, holding all the other variables constant when the individual is male,the individual purchase amount increase 29.00%; holding all the other variables constant when the individual is in city C,the individual purchase amount decrease 58.80%. At 99.9% significance level, holding all the other variable constant, when the individual is in city B, the individual purchase amount increase 11.34%. At 99% significance level, holding all the other variable constant, when the individual is in occupation 7, the individual purchase amount decrease 10.70%.

d. From the results of individual and multiple nonlinear regression, we found these variables (gender, city_Category) significantly influence the percentage change in purchase. Men have a significantly influence in  increasing purchase amount. So from this we can predict, men have a large influence in increasing purchasing power. Individual living in city C has a significanly influence decreasing in purchase amount. So from this we can predict that individual in city has significantly decreasing purchasing power (or no puchasing power).

e. The omittied variable biaas is the variable which we didn't count in the above analysis which can influence dependent variable log(purchase) and is also correlated with the independent variables. So the ommited variabel can be "annual salary" which can influence individul in changing in purchase, and also is correlated with occupation. The other omitted varibel can be "ethnicity", which is also influence the dependent variable, and is also correlated with occupation or city_category. However, we don't have enough data for these variables, which we can't do regression on this. Moreover, the above regresion model is showing the effect on log purchase of one independet variable, holding other variables constant. The limitation of this is the effect of marital status in this specification, holding constnat gender, is the same for men and women. However, no reason that this must be so. The effect on percentage change in purchase amount of marital stautus holding gender constant, could depende on the difference of the gender. So the interaction term is also an example of control variable. The definition of control variable is that it is correlated with omitted causal factor, but it does not necessarily have a cauasl effect on dependent variable. We decide to use put interaction term of gender and marital_stauts, the interaction term of city category and years staying in the city in our new regression equation as control variables.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
lm_omitted=lm(log(Purchase)~Gender+Age+Occupation+City_Category+Stay_In_Current_City_Years+Marital_Status
+Gender*Marital_Status+City_Category*Stay_In_Current_City_Years,data=demo)

summary(lm_omitted)

```

Explanation: We run a regression of log purchase on all the independent variables and the control variables together. We find at 99.99% significance level, holding all the other variables constant when the individual is male,the individual purchase amount increase 30.73%; holding all the other variables constant when the individual is in city C,the individual purchase amount decrease 72.20%. At 99.9% significance level, holding all the other variable constant, when the individual is in age 26-35, the individual purchase amount increase 27.38%. At 99% significance level, holding all the other variable constant, when the individual is living in city C and stay in more than 4 years, the individual purchase amount increase 12.08%. The R sqare change from 0.1514 to 0.1542. So from this, the higher R square means that the regressors with control variables explain the variation in log(purchase) bettwen than just without. From both regression equations, we interpret that male individual has signifcantly increasing purchaing power, and city C has significantly devreasing purchasing power.

**6. Segmentation**
```{r}
data = model.matrix(~Gender+Age+Occupation+City_Category+Marital_Status+Stay_In_Current_City_Years,data = demo)

scaled_data = scale(data,scale=FALSE)
k.max <- 15
wss <- sapply(1:k.max, 
              function(k){kmeans(scaled_data, k, nstart=50,iter.max = 20 )$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")

seg.summ <- function(data, groups) {
  aggregate(data, list(groups), function(x) mean(as.numeric(x))) 
}  
set.seed(96743)        
seg.k3 <- kmeans(scaled_data, centers=3)

seg.summ(data, seg.k3$cluster)

seg.k4 <- kmeans(scaled_data, centers=4)

seg.summ(data, seg.k4$cluster)
```

Explanation 

```{r}
demo$Cluster = seg.k4$cluster
demo
```

Explanation

**Conclusion**

From our preliminary data analysis, we were able to determine customer demographics that were associated with higher purchase amounts. In particular, we found that men spend more money on Black Friday than women, the purchase amounts are higher in different cities (City C has higher purchase amounts), unmarried people spend more than married (but only slightly more), certain occupation categories spend more than others, and interestingly, that there was not a huge difference in mean purchase by age (26-35 category spends slightly more). Also interestingly, there was almost no perceivable difference in mean purchase by years in city. For the marital status, its effect on purchase was ambiguous. Based on the current data, we cannot confidentially say that there exists difference between married and unmarried customers. We need more data to analyze this problem later.

By using statistical tests and visualizations, we were able to easily pinpoint particular demographics on which the company should direct its marketing dollars. Further analysis will illuminate these important demographics.

[^1]: American shoppers spent a record $5 billion in 24 hours. That marks a 16.9% increase in dollars spent online compared with Black Friday 2016. “Online Sales Boomed on Black Friday.” CNNMoney, Cable News Network, money.cnn.com/2017/11/25/news/black-friday-holiday-shopping-foot-traffic/index.html.