---
title: "Interim Report"
author: "GZ"
date: "10/8/2018"
output: html_document
---
#Final Project: Interim Repot 1
##Quenstion 2 
###a)
```{r, echo=FALSE}
library(rmarkdown)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)

setwd("~/Desktop/marketing analytics")
 BF <- read_csv("~/Desktop/marketing analytics/BlackFriday.csv")
 
 str(BF)
 dim(BF)

```
*Answer*
Measure types of conceptual variables: Character-"Product_ID","Gender","Age","Stay_In_Current_City_Years"; 
Integer-"Occupation","Marital_Status","Product_Category_1","Product_Category_2","Product_Category_3","Purchase"

```{r}
#check if data cotains missing values
install.packages("VIM")
library('VIM')
library(dplyr)
data(BF, package = "VIM")
dim(BF)
sum(complete.cases(BF))
aggr(BF,cex.axis = .4)

#set missing value as "0"
BF$Product_Category_2[is.na(BF$Product_Category_2)]<-"0"
BF$Product_Category_3[is.na(BF$Product_Category_3)]<-"0"


#clean the data
demo<-BF %>% 
select(-starts_with("Product")) %>% 
group_by(User_ID) %>% 
mutate(Purchase = sum(Purchase)) %>% 
distinct(User_ID,.keep_all=TRUE)

#coerce the data type
demo$Gender<-as.factor(demo$Gender)
demo$Age<-as.factor(demo$Age)
demo$Occupation<-as.factor(demo$Occupation)
demo$City_Category<-as.factor(demo$City_Category)
demo$Stay_In_Current_City_Years<-as.factor(demo$Stay_In_Current_City_Years)
demo$Marital_Status<-as.factor(demo$Marital_Status)

glimpse(demo)
```

## Question 3

```{r}

### I want to see the purchase power of different gender. I choose variable gender, and averge_purchase
ggplot(demo, aes(x=Gender, y = Purchase)) + geom_boxplot()+theme(plot.title = element_text(hjust = 0.5)) +labs(title="Individual Purchase in different gender", x = "Gender", y = "purchase")

demo_gender<-demo %>% 
  group_by(Gender) %>% 
  summarise(count= n(),
            average_purchase=sum(as.numeric(Purchase))/count)
            
  
ggplot(demo_gender, aes(x=`Gender`,y=`average_purchase`,fill=`Gender`)) +
  geom_bar(stat="identity",width=0.7,position = "dodge") +
  geom_text(aes(label=count),position = position_dodge(width = 1),vjust =-0.5) +
  labs(title="Purchasing power in different gender", x = "Gender", y = "Average_purchase") +
  theme(plot.title = element_text(hjust = 0.5))

### I want to se purchasing power in different age level
ggplot(demo, aes(x=Age, y = Purchase)) + geom_boxplot()+theme(plot.title = element_text(hjust = 0.5)) +labs(title="Individual Purchase in different age", x = "Age", y = "purchase")

demo_age<-demo %>% 
  group_by(Age) %>% 
  summarise(count=n(),average_purchase=sum(as.numeric(Purchase))/count)

ggplot(demo_age, aes(x=`Age`,y=`average_purchase`,fill=`Age`)) +
  geom_bar(stat="identity",width=0.7,position = "dodge") +
  geom_text(aes(label=count),position = position_dodge(width = 1),vjust =-0.5) +
  labs(title="Purchasing power in different age", x = "Age", y = "Average_purchase") +
  theme(plot.title = element_text(hjust = 0.5))

### I want to see purchasing power in different occupation
demo_occupation<-demo %>% 
  group_by(Occupation) %>% 
  summarise(count=n(),average_purchase=sum(as.numeric(Purchase))/count)

ggplot(demo_occupation, aes(x=`Occupation`,y=`average_purchase`,fill=`Occupation`)) +
  geom_bar(stat="identity",width=0.7,position = "dodge") +
  geom_text(aes(label=count),position = position_dodge(width = 1),vjust =-0.5) +
  labs(title="Purchasing power in different occupation", x = "Occpation", y = "Average_purchase") +
  theme(plot.title = element_text(hjust = 0.5))

### I want to see purchasing power in city category and city year
demo_city<-demo %>% 
  group_by(City_Category, Stay_In_Current_City_Years) %>% 
  summarise(count=n(),average_purchase=sum(as.numeric(Purchase))/count)

ggplot(demo_city, aes(x=`City_Category`,y=`average_purchase`,fill=`City_Category`)) +
  geom_bar(stat="identity",width=0.7,position = "dodge") +
  geom_text(aes(label=count),position = position_dodge(width = 1),vjust =-0.5) +
  facet_wrap(~Stay_In_Current_City_Years)+
  labs(title="Purchasing power in different city", x = "City", y = "Average_purchase") +
  theme(plot.title = element_text(hjust = 0.5))

### I want to see purchasing power in marriage status
ggplot(demo, aes(x=Marital_Status, y = Purchase)) + geom_boxplot()+theme(plot.title = element_text(hjust = 0.5)) +labs(title="Individual Purchase in different marriage statue", x = "Marital_Statue", y = "purchase")

demo_marriage_status<-demo %>% 
  group_by(Marital_Status) %>% 
  summarise(count=n(),average_purchase=sum(as.numeric(Purchase))/count)
  
ggplot(demo_marriage_status, aes(x=`Marital_Status`,y=`average_purchase`,fill=`Marital_Status`)) +
  geom_bar(stat="identity",width=0.7,position = "dodge") +
  geom_text(aes(label=count),position = position_dodge(width = 1),vjust =-0.5) +
  labs(title="Purchasing power in different marital status", x = "Marital Status", y = "Average_purchase") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
